Cluster Dynamics with HETDEX at z < 0.5 - I: Simulated Performance, Mass Recovery, and Limits to Cosmology

### 2016-07-11 paper1.tex
Eduardo's comments

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
%\usepackage{epsf}
\usepackage{color}
\usepackage{graphicx}
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{mathrsfs}	% Extra extra math symbols
\usepackage{natbib}
%\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{multirow}
\usepackage{etoolbox}
\usepackage{listings}

@@

\begin{abstract}
\noindent
The distribution of massive clusters of galaxies depends strongly on the total cosmic mass density, the mass variance, and the dark energy equation of state. As such, measures of galaxy clusters can provide constraints on these parameters and even test models of gravity, but only if observations of clusters can lead to accurate estimates of their total masses. Here, we carry out a study to investigate the ability of a blind spectroscopic survey to recover accurate galaxy cluster masses through their velocity dispersion using probability based and machine learning methods. We focus on the Hobby Eberly Telescope Dark Energy Experiment (HETDEX), which will employ new Visible Integral-Field Replicable Unit Spectrographs (VIRUS), over 420 \degsq\ on the sky with a 1/4.5 fill factor. VIRUS covers the blue/optical portion of the spectrum ($3500-5500~\AAA$), allowing surveys to measure redshifts for a large sample of galaxies out to $z < 0.5$ based on their absorption features or [\ion{O}{ii}] $\lambda$3727 emission (and Lyman-$\alpha$ over $1.9 < z < 3.5$). We use a detailed mock galaxy catalog from a semi-analytic model to simulate surveys observed with VIRUS, including: (1) a blind, HETDEX-like survey with an incomplete but uniform spectroscopic selection function; and (2) IFU surveys that target clusters directly, obtaining spectra of all galaxies in a VIRUS-sized field. For both surveys, we include realistic uncertainties from galaxy magnitude and line-flux limits. We benchmark both surveys against spectroscopic observations with ``perfect" knowledge of galaxy line-of-sight velocities. With HETDEX, we can recover cluster masses to $\sim0.1$ dex which can be further improved to $<0.1$ dex with targeted follow-up observations. This level of cluster mass recovery enables independent constraints on $\sigma_8$ to $<20$\%, and the unique properties of the observations will provide important calibrations for the zeropoint and scatter in the optical richness-mass relation.
\end{abstract}

@@

As mass is not a direct observable, a lot of work is underway to characterize galaxy cluster masses with an observable feature of galaxy clusters \citeeg{Mantz2016,Bocquet2015,Old2015,Old2014, Pearson2015}. The goal is to constrain $P(M|z, \vec{x})$ the probability density ($P$) that a galaxy cluster of given mass ($M$), located at a specific redshift ($z$), determined using an observable parameter or parameters ($\vec{x}$). Generally, cluster mass calibrations are done in one of two ways, through simulations or direct or statistical calibration.

@@

The second broad method is the direct calibration of cluster masses. This recipe has two distinct but not always independent tracks. The ``direct'' method uses observations of a relatively small set of clusters and then uses known mass estimators, including X-ray temperatures and luminosities \citeeg{Mantz2010, Rykoff2014}, microwave observations \citeeg{Vanderlinde2010, Sehgal2011}, optical richness \citeeg{Abell1958, Rykoff2012} or weak lensing (WL; \eg\ \citealt{Rozo2010}) as examples, which provide a ``true'' mass. This directly calibrates the observable-mass relation which is then applied to a much larger sample. The complications lie in that the ``true'' masses are, in fact, estimations, and the methods used to recover these cluster masses are subject to their own limitations. X-ray based cluster masses assume hydrostatic equilibrium \citeeg{Mantz2015} which may only be valid for a very small number and range of cluster masses. The Sunyaev-Zel'dovich Effect (SZE; \citealt{Sunyaev1972}), which uses the up-scattering of cosmic microwave background (CMB) photons to estimate cluster masses, provides accurate estimations of mass, but the ability to detect low mass galaxy clusters is currently limited by technology \citeeg{Carlstrom2002a} and can also be effected by the properties of the intracluster medium \citeeg{Pipino2010}. WL estimates are, in principle, correct in the mean, but they suffer from signal-to-noise requirements, limiting their usefulness in low mass clusters (where the lensing signal is particularly weak), and potentially suffer from line-of-sight effects as WL is sensitive to all mass along the line-of-sight. Virial mass estimators which determine the cluster mass based on the motions of the member galaxies \citeeg{Ruel2014, Sifon2015} are promising in that it is a direct measurement of the depth of clusters potential well, but suffer from systematics due to cluster formation physics which disrupts the velocity field.

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{figures/oii_sdss.pdf} 
	\caption[Illustration of the probability based \hbox{[\ion{O}{II}]} flux prediction method.]{\textit{Left}: CMD of 503,113 $z<0.2$ galaxies take from the SDSS DR12 where the shading scales with the density of points. The two colored boxes show regions containing potential catalog galaxies. \textit{Right}: Probability histograms of the Log \hbox{[\ion{O}{II}]} luminosity for the SDSS galaxies located in the two highlighted regions on the right. The \hbox{[\ion{O}{II}]} luminosities are assigned to catalog galaxies from slice sampling the probability histogram and converted to fluxes using the redshift of each galaxy.}
	\label{fig: oii sdss} 
\end{figure*}

@@

We begin in Section \ref{sec:Data} by giving an overview of what data is used, how it is created, and how we make our ``observations.'' Details about the determination of cluster parameters, velocity dispersion, total mass, etc., are discussed in Section \ref{sec:recovery}. Next, we present the results of our study in Section \ref{sec:results} and discuss their implications in Section \ref{sec:discussion}. Finally we summarize our findings in Section \ref{sec:summary}. A follow-up to this work Boada \etal\ (in Preparation) will investigate how the techniques developed here will work in practice.

@@

The catalog information, used in this study, is broken into two large portions. The ``truth'' files contain the characteristics of each individual galaxy, such as right ascension (RA), declination (DEC), redshift ($z$), observed and rest-frame magnitudes, and many others. The ``halo'' files contain information for individual halos, to which many individual galaxies may belong. This includes five estimations of dynamical mass, RA, DEC, z, three dimensional velocity dispersion, and many others. However, the catalogs do not include information for emission lines. We supplement the catalogs by generating this information; the process is described in Section~\ref{sec: oii luminosity}.

@@

We investigate the accuracy of the halo mass distribution by comparing the cumulative number density of halos above a mass ($M_{200c}$) threshold to the halo mass function (HMF) of \cite{Tinker2008}. We calculate the HMF at central redshifts of 0.1, 0.2, and 0.4 using {\sc HMFcalc} \citep{Murray2013} and compare it galaxies in a redshift window of $\Delta z\pm0.01$. We find a very good agreement between the predicted HMF and the observed distribution of clusters.

\subsection{Conditional {\rm[\ion{O}{II}]} Flux Probability Distribution Functions}\label{sec: oii luminosity}
We use the SDSS DR12 \citep{Alam2015} catalogs to assign [\ion{O}{II}] emission line strengths to the galaxies in the Buzzard catalog. We use 503,113 objects classified as galaxies selected over $z = 0.02 - 0.2$ with {\sc zwarning $=0$} and a measured [\ion{O}{II}] line flux signal-to-noise of five. Figure~\ref{fig: oii sdss} shows the color-magnitude diagram (CMD) of $M_r$ and $g-r$ for these galaxies.

@@

Figure~\ref{fig: oii sdss} illustrates this process. The numbered boxes in the left panel show the bins corresponding to two example Buzzard galaxies ($M_r, g-r = -17.7,~0.49$ and $M_r, g-r = -21.4,~1.24$). The right panel shows the Log [\ion{O}{II}] luminosity distribution functions, $P($[\ion{O}{II}]$| M_r, g-r)$, which we use to assign [\ion{O}{II}] luminosity to each object. The luminosity is then converted into an [\ion{O}{II}] flux through 
\begin{equation}
	F = \frac{L}{4\pi D_L}
\end{equation}
where $D_L$ is the luminosity distance \citeeg{Hogg1999}.

@@

In simple terms, the cluster redshift is the mean of the redshifts of all galaxies associated with the cluster, where the mean is the first moment of the velocity (redshift) distribution function $P(z)$. In practice, the first moment is strongly subject to outliers, so we rely instead on the biweight location estimator \citep{Beers1990} through\footnote{Implemented as part of the {\sc astLib} Python library. See \hbox{http://astlib.sourceforge.net}}:
\begin{equation}
z_c = z_{med} + \frac{\sum_{|u_i| < 1} (z_i - M)(1-u_i^2)^2}{\sum_{|u_i| < 1}(1-u_i^2)^2}
\end{equation}
where $z_i$ are the individual redshifts, $z_{med}$ is the median redshift and $u_i$ is given by:
\begin{equation}
	u_i = \frac{(z_i - M)}{C\,\mathrm{MAD}}.
\end{equation}
MAD is the median absolute deviation, also defined in \cite{Beers1990}, and $C$ is the a tuning constant. We choose $C=6$ (the suggested value) which balances computational speed and location accuracy. 
% Although this work assumes that we know each galaxy's redshift to infinite precision, in practice, we find a simple weighted mean provides a reliable estimate of $z_c$ when there are uncertainties on the individual galaxy redshifts.

@@

We follow the maximum likelihood method of \cite{Walker2006} to estimate the line-of-sight velocity dispersion (LOSVD). Using the previously calculated LOSVs ($v_i$ in the following discussion), we maximize the probability function:
\begin{equation}\label{eq: jointGaussian}
P(\{v_1, ..., v_N\})=\displaystyle\prod_{i=1}^{N}\frac{1}{\sqrt{2\pi(\sigma_i^2+\sigma_{1D}^2)}}\exp\biggl[-\frac{1}{2}\frac{(v_i-\langle u \rangle)^2}{(\sigma_i^2+\sigma_{1D}^2)}\biggr]
\end{equation}
where $\sigma_{1D}$, $\langle u \rangle$, and $\sigma_i$ is the LOSVD, the average radial velocity and the error (which we have assumed to be normally distributed) on the individual $v_i$ measurements respectively. To maximize Equation~\ref{eq: jointGaussian}, we employ a Monte Carlo Markov Chain (MCMC) sampler ({\sc emcee}\footnote{http://dan.iel.fm/emcee/current/}; \citealt{Foreman-Mackey2013}) which is based on affine-invariant ensemble sampler (see \citealt{Goodman2010} for details on affine-invariant samplers). We draw twenty thousand samples from the posterior probability distribution using simple priors, $\langle\mu\rangle$ lies between the maximum and minimum LOSV and $0< \sigma_{1D} < 1400$ \kms, for each cluster. We set the upper limit on the LOSVD as the LOSVD corresponding to a $10^{16}$ \Msol\ cluster at $z\sim0.0$, higher mass than any expected cluster in the Buzzard catalogs. When the complete sampling of the posterior probability distribution is not used, the final LOSVD is quoted as the median value of the posterior probability distribution with 68\% error bars defined as the square root of second moment of the same distribution, the standard deviation.

@@

A growing body of work suggests that there is a significant difference in the observed LOSVD depending on the velocity tracers used. Specifically, while there is little difference between using galaxies and their host DM subhalos, there is a significant over estimation of the LOSVD when using galaxies/subhalos compared to DM particles \citep{Munari2013}. We follow other works \citeeg{Kirk2015, Sifon2015a} using the scaling relation, given in Equation~\ref{eq:power law} to facilitate comparisons with other observational studies, which rely on galaxies as tracers.

@@

\subsubsection{Probability Based}\label{sec:probability method}
For internal consistency between this and the ML based method we use 70\% of the clusters to establish a conditional probability density of $P(M_{200c} | \vec{x})$ which we then use as the cluster mass probability density for the remaining clusters with similar features. In this way we, ``train'' the probability density using the existing simulated data, and apply it to the ``test'' sample (the remaining 30\% of the data not used as the training set).

\begin{figure*}
		\includegraphics[width=\textwidth]{figures/prob_example.pdf} 
	\caption[Illustration of the probability based cluster mass prediction method.]{Illustration of the probability based cluster mass prediction method. \emph{Left}: The two dimensional posterior probability distribution of LOSVD and redshift used to determine the correct cluster mass. The pink and green rectangles show the locations of two example galaxies used to create the conditional probability distribution of the mass, $P(M_{200c}|z, \mathrm{Log}\, \sigma)$. \emph{Right}: The conditional probability distribution of the cluster mass for the two example galaxies. See text for a complete description.}
	\label{fig: probability corner} 
\end{figure*}

In this method, the cluster masses are predicted using the method illustrated in Figure~\ref{fig: probability corner}. The left panel shows the two dimensional (joint probability) projections of the posterior probability distributions of the feature training data. The conditional probability of the cluster mass $P(M_{200c}|\vec{x}= \{ x_1,x_2,...\})$ (shown in the right panel) is determined by selecting a region in the joint probability distribution. For example, using the LOSVD and redshift features we create $P(M_{200c}|\vec{x})$ for two test galaxies, shown by the green and pink boxes in the left panel of Figure~\ref{fig: probability corner}. These example galaxies have features $\vec{x} = \{\sigma=200\, \kms,z=0.16\}$ and $\vec{x} = \{\sigma=400\, \kms,z=0.36\}$. We select all galaxies in our training sample with similar features and create the conditional probability distributions shown in the right panel.

For the clusters making up the \emph{test} sample the mass is unknown (it is what we are trying to predict) but the other features are known. To determine the mass probability distribution of a test cluster, $P(M_{200c})$, we combine the conditional probability distribution, $P(M_{200c}|\vec{x})$, created previously with the probability distribution of $\sigma$, the LOSVD, through Equation~\ref{eq: Pm}.
\begin{equation}\label{eq: Pm}
	P(M_{200c}) = \int P(M_{200c}|\vec{x})\ P(\sigma)\ d\sigma\ P(z)\ dz
\end{equation}
The expected mass is determined by calculating the first moment of the probability density. This becomes our ``predicted'' cluster mass, $M_{pred}$.
\begin{equation}\label{eq: expected mass}
	M_{pred}= \int M_{200c}\ P(M_{200c})\ dM_{200c}
\end{equation}
The confidence interval associated with this prediction can be estimated two ways. First, by calculating the second moment of the probability density through
\begin{equation}\label{eq: variance}
	\mathrm{Var} = \int (M_{200c} - M_{pred})^2\ P(M_{200c})\ dM_{200c}
\end{equation}
or by drawing many samples from $P(M_{200c})$ and calculating the values at the 16th and 84th percentile. In practice we find that both methods produce similar results for a large number of trials. Therefore, we quote predicted masses as the most probable mass given by Equation~\ref{eq: expected mass} and associated 68\% error estimated through the square root of Equation~\ref{eq: variance}.

\subsubsection{Machine Learning Based}\label{sec:machine learning method}
The cluster mass estimation in this section relies on a ML technique known as an ensemble method, where many estimators are created by a single learning method with the goal of improved generalization and robustness compared to a single estimation. Ensemble methods \citeeg{Caruana2006} come in two general flavors. Averaging methods average (hence the name) the estimators to produce a single prediction. Boosting estimators build estimates sequentially by attempting to address poor performing estimators in each previous step, hence ``boosting'' the predictive power.

@@

As discussed in Section \ref{sec: observations}, the observational constraints place limits on the total number of clusters member galaxies expected to be recovered. Knowing these limits will provide important information for potential future follow up or Targeted observations. We recover 14,189 clusters with Targeted observations and 1,760 clusters with Survey observations, where we require a detection of $N_{obs} >=5$ galaxies for a cluster to be detected. Figure \ref{fig: recovery} shows the recovery fraction of member galaxies, the number of observed galaxies divided by the number of actual galaxies ($N_{obs}/N_{True}$), as a function of both redshift and cluster mass. As expected, the Targeted observing strategy where the individual clusters are targeted through several dithers to ensure near complete coverage, performs significantly better than the Survey observing strategy across all redshifts and cluster masses. By definition, with Perfect knowledge, recovery fraction would be unity across all redshifts and cluster masses where clusters exist.

@@

\subsection{Mass estimates}
\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/Probcomparison.pdf} 
	\caption[Probability based cluster mass predictions.]{Mass predictions for the power law scaling relation (Equation~\ref{eq:power law}) and the probability based technique with different input features as a function of true cluster mass. The bottom row of panels shows the fractional error (Equation~\ref{eq: fractional error}) also as a function of true cluster mass. The solid black line shows the 1:1 relation, in both panels, between $M_{200c}$ and $M_{pred}$. The orange, dash-dotted line is the median predicted mass for perfect observations. The solid, purple line is the median predicted mass for the Targeted observing, and the green, dashed line is the median recovered mass for the HETDEX-like observations. The shaded regions represent the 68\% scatter around the median values.}
	\label{fig:Probability comparison} 
\end{figure*}

@@

\subsection{Impact of Training Sample Cosmology}
All simulations (including that used for Buzzard) use specific values for cosmological parameters. When using simulation data to train ML methods, we incorporate all of those assumptions into the learned feature associations. One could imagine that the specific values of the cosmological parameters in the training sample could bias the ML results when applied to data (real or simulated) created from an unknown true set of cosmological parameters.

>>>>

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
%\usepackage{epsf}
\usepackage{color}
\usepackage{graphicx}
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{mathrsfs}	% Extra extra math symbols
\usepackage{natbib}
\usepackage[colorlinks,urlcolor=magenta,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{multirow}
\usepackage{etoolbox}
\usepackage{listings}

@@

\begin{abstract}
\noindent
The distribution of massive clusters of galaxies depends strongly on the total cosmic mass density, the mass variance, and the dark energy equation of state. As such, measures of galaxy clusters can provide constraints on these parameters and even test models of gravity, but only if observations of clusters can lead to accurate estimates of their total masses. Here, we carry out a study to investigate the ability of a blind spectroscopic survey to recover accurate galaxy cluster masses through their velocity dispersion using probability based and machine learning methods. We focus on the Hobby Eberly Telescope Dark Energy Experiment (HETDEX), which will employ new Visible Integral-Field Replicable Unit Spectrographs (VIRUS), over 420 \degsq\ on the sky with a 1/4.5 fill factor. VIRUS covers the blue/optical portion of the spectrum ($3500-5500~\AAA$), allowing surveys to measure redshifts for a large sample of galaxies out to $z < 0.5$ based on their absorption features or [\ion{O}{ii}] $\lambda$3727 emission (and Lyman-$\alpha$ over $1.9 < z < 3.5$). We use a detailed mock galaxy catalog from a semi-analytic model to simulate surveys observed with VIRUS, including: (1) a blind, HETDEX-like survey with an incomplete but uniform spectroscopic selection function; and (2) integral field unit surveys that target clusters directly, obtaining spectra of all galaxies in a VIRUS-sized field. For both surveys, we include realistic uncertainties from galaxy magnitude and line-flux limits. We benchmark both surveys against spectroscopic observations with ``perfect" knowledge of galaxy line-of-sight velocities. With HETDEX, we can recover cluster masses to $\sim0.1$ dex which can be further improved to $<0.1$ dex with targeted follow-up observations. This level of cluster mass recovery enables independent constraints on $\sigma_8$ to $<20$\%, and the unique properties of the observations will provide important calibrations for the zeropoint and scatter in the optical richness--mass relation.
\end{abstract}

@@

As mass is not a direct observable, a lot of work is underway to characterize galaxy cluster masses with an observable feature of galaxy clusters \citeeg{Mantz2016,Bocquet2015,Old2015,Old2014, Pearson2015}. The goal is to constrain $P(M|z, \vec{x})$ the probability density ($P$) that a galaxy cluster of given mass ($M$), located at a specific redshift ($z$), determined using an observable parameter or parameters ($\vec{x}$). Generally, cluster mass calibrations are done in one of three ways, through simulations, direct or statistical calibration.

@@

The second broad method is the direct calibration of cluster masses. This recipe has two distinct but not always independent tracks. The ``direct'' method uses observations of a relatively small set of clusters and then uses observable mass proxies, including X-ray temperatures and luminosities \citeeg{Mantz2010, Rykoff2014}, microwave observations \citeeg{Vanderlinde2010, Sehgal2011}, optical richness \citeeg{Abell1958, Rykoff2012} or weak lensing (WL; \eg\ \citealt{Rozo2010}) as examples, which provide a ``true'' mass. This directly calibrates the observable-mass relation which is then applied to a much larger sample. The complications lie in that the ``true'' masses are, in fact, estimations, and the methods used to recover these cluster masses are subject to their own limitations. X-ray based cluster masses assume hydrostatic equilibrium \citeeg{Mantz2015} which may only be valid for a very small number and range of cluster masses. The Sunyaev-Zel'dovich Effect (SZE; \citealt{Sunyaev1972}), which uses the up-scattering of cosmic microwave background (CMB) photons to estimate cluster masses, provides accurate estimations of mass, but the ability to detect low mass galaxy clusters is currently limited by technology \citeeg{Carlstrom2002a} and can also be effected by the properties of the intracluster medium \citeeg{Pipino2010}. WL estimates are, in principle, correct in the mean, but they suffer from signal-to-noise requirements, limiting their usefulness in low mass clusters (where the lensing signal is particularly weak), and potentially suffer from line-of-sight effects as WL is sensitive to all mass along the line-of-sight. Virial mass estimators which determine the cluster mass based on the motions of the member galaxies \citeeg{Ruel2014, Sifon2015} are promising in that it is a direct measurement of the depth of clusters potential well, but suffer from systematics due to cluster formation physics which disrupts the velocity field.

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{figures/oii_sdss.pdf} 
	\caption{\textit{Left}: CMD of 503,113 $z<0.2$ galaxies take from the SDSS DR12 where the shading scales with the density of points. The two colored boxes show regions containing potential catalog galaxies. \textit{Right}: Probability histograms of the Log \hbox{[\ion{O}{II}]} luminosity for the SDSS galaxies located in the two highlighted regions on the right. The \hbox{[\ion{O}{II}]} luminosities are assigned to catalog galaxies from slice sampling the probability histogram and converted to fluxes using the redshift of each galaxy.}
	\label{fig: oii sdss} 
\end{figure*}

@@

We begin in Section \ref{sec:Data} by giving an overview of what data is used, how it is created, and how we make our ``observations.'' Details about the determination of cluster parameters, velocity dispersion, total mass, etc., are discussed in Section \ref{sec:recovery}. Next, we present the results of our study in Section \ref{sec:results} and discuss their implications in Section \ref{sec:discussion}. Finally we summarize our findings in Section \ref{sec:summary}. A follow-up to this work Boada \etal\ (in Preparation) will will apply these techniques to real data so as to empirically validate our conclusions.

@@

The catalog information used in this study is broken into two large portions. The ``truth'' files contain the characteristics of each individual galaxy, such as right ascension (RA), declination (DEC), redshift ($z$), observed and rest-frame magnitudes, and many others. The ``halo'' files contain information for individual halos, to which many individual galaxies may belong. This includes estimations of dynamical mass, RA, DEC, z, three dimensional velocity dispersion, and many others. However, the catalogs do not include information for emission lines. We supplement the catalogs by generating this information; the process is described in Section~\ref{sec: oii luminosity}.

@@

We investigate the accuracy of the halo mass distribution by comparing the cumulative number density of halos above a mass ($M_{200c}$) threshold to the halo mass function (HMF) of \cite{Tinker2008}. We calculate the HMF at central redshifts of 0.1, 0.2, and 0.4 using {\sc HMFcalc} \citep{Murray2013} and compare it to galaxy clusters in a redshift window of $\Delta z\pm0.01$. We find a very good agreement between the predicted HMF and the observed distribution of clusters.

\subsection{Conditional {\rm[\ion{O}{II}]} Flux Probability Distribution Functions}\label{sec: oii luminosity}
We use the SDSS DR12 \citep{Alam2015} catalogs to assign [\ion{O}{II}] emission line strengths to the galaxies in the Buzzard catalog. We use 503,113 objects classified as galaxies selected over $z = 0.02 - 0.2$ with {\sc zwarning $=0$} and a measured [\ion{O}{II}] line flux signal-to-noise of five. Figure~\ref{fig: oii sdss} shows the color-magnitude diagram (CMD) of $M_r$ and $g-r$ for these galaxies, and the following SQL query selects the appropriate galaxies:
\begin{lstlisting}[language=SQL]
SELECT GAL.objid, 
       GAL.u, 
       GAL.err_u, 
       GAL.g, 
       GAl.err_g, 
       GAL.r, 
       GAL.err_r, 
       GAL.i, 
       GAl.err_i, 
       GAL.z, 
       GAL.err_z, 
       g.oii_3726_flux, 
       g.oii_3726_flux_err, 
       s.z 
FROM   galaxy AS GAL 
       JOIN specobj AS s 
         ON s.bestobjid = GAL.objid 
       JOIN galspecline AS g 
         ON g.specobjid = s.specobjid 
WHERE  oii_3726_flux > oii_3726_flux_err * 5 
       AND oii_3726_flux_err > 0 
       AND s.class = 'GALAXY' 
       AND s.zwarning = 0 
\end{lstlisting}

@@

Figure~\ref{fig: oii sdss} illustrates this process. The numbered boxes in the left panel show the bins corresponding to two example Buzzard galaxies ($M_r, g-r = -17.7,~0.49$ and $M_r, g-r = -21.4,~1.24$). The right panel shows the [\ion{O}{II}] luminosity distribution functions, $P(\log\ L_{[\ion{O}{II}]}| M_r, g-r)$, which we use to assign [\ion{O}{II}] luminosity to each object. The luminosity is then converted into an [\ion{O}{II}] flux through 
\begin{equation}
	F = \frac{L}{4\pi D_L^2}
\end{equation}
where $D_L$ is the luminosity distance \citeeg{Hogg1999}.

@@

In simple terms, the cluster redshift is the mean of the redshifts of all galaxies associated with the cluster, where the mean is the first moment of the velocity (redshift) distribution function $P(z)$. In practice, the first moment is strongly subject to outliers, so we rely instead on the biweight location estimator \citep{Beers1990} through\footnote{Implemented as part of the {\sc astLib} Python library. See \hbox{\url{http://astlib.sourceforge.net}}}:
\begin{equation}
z_c = z_{med} + \frac{\sum_{|u_i| < 1} (z_i - M)(1-u_i^2)^2}{\sum_{|u_i| < 1}(1-u_i^2)^2}
\end{equation}
where $z_i$ are the individual redshifts, $z_{med}$ is the median redshift and $u_i$ is given by:
\begin{equation}
	u_i = \frac{(z_i - M)}{C\,\mathrm{MAD}}.
\end{equation}
MAD is the median absolute deviation, also defined in \cite{Beers1990}, and $C$ is the a tuning constant. We choose $C=6$ (the suggested value) which balances computational speed and location accuracy. 
% Although this work assumes that we know each galaxy's redshift to infinite precision, in practice, we find a simple weighted mean provides a reliable estimate of $z_c$ when there are uncertainties on the individual galaxy redshifts.

@@

We follow the maximum likelihood method of \cite{Walker2006} to estimate the line-of-sight velocity dispersion (LOSVD). Using the previously calculated LOSVs ($v_i$ in the following discussion), we maximize the probability function:
\begin{equation}\label{eq: jointGaussian}
P(\{v_1, ..., v_N\})=\displaystyle\prod_{i=1}^{N}\frac{1}{\sqrt{2\pi(\sigma_i^2+\sigma_{1D}^2)}}\exp\biggl[-\frac{1}{2}\frac{(v_i-\langle u \rangle)^2}{(\sigma_i^2+\sigma_{1D}^2)}\biggr]
\end{equation}
	where $\sigma_{1D}$, $\langle u \rangle$, and $\sigma_i$ is the LOSVD, the average radial velocity and the error (which we have assumed to be normally distributed) on the individual $v_i$ measurements respectively. To maximize Equation~\ref{eq: jointGaussian}, we employ a Monte Carlo Markov Chain (MCMC) sampler ({\sc emcee}\footnote{\url{http://dan.iel.fm/emcee/current/}}; \citealt{Foreman-Mackey2013}) which is based on affine-invariant ensemble sampler (see \citealt{Goodman2010} for details on affine-invariant samplers). We draw twenty thousand samples from the posterior probability distribution using simple priors, $\langle\mu\rangle$ lies between the maximum and minimum LOSV and $0< \sigma_{1D} < 1400$ \kms, for each cluster. We set the upper limit on the LOSVD as the LOSVD corresponding to a $10^{16}$ \Msol\ cluster at $z\sim0.0$, higher mass than any expected cluster in the Buzzard catalogs. When the complete sampling of the posterior probability distribution is not used, the final LOSVD is quoted as the median value of the posterior probability distribution with 68\% error bars defined as the square root of second moment of the same distribution, the standard deviation.

@@

Recent work suggests that there is a significant difference in the observed LOSVD depending on the velocity tracers used \citep{Munari2013}. Specifically, while there is little difference between using galaxies and their host DM subhalos, there is a significant over estimation of the LOSVD when using galaxies/subhalos compared to DM particles. We follow other works \citeeg{Kirk2015, Sifon2015a} using the scaling relation, given in Equation~\ref{eq:power law} to facilitate comparisons with other observational studies, which rely on galaxies as tracers.

@@

\subsubsection{Probability Based}\label{sec:probability method}
For internal consistency between this and the ML based method we use 70\% of the clusters to establish a conditional probability density of $P(\log\ M_{200c} | \vec{x})$ which we then use as the cluster mass probability density for the remaining clusters with similar features. In this way we, ``train'' the probability density using the existing simulated data, and apply it to the ``test'' sample (the remaining 30\% of the data not used as the training set).

\begin{figure*}
		\includegraphics[width=\textwidth]{figures/prob_example.pdf} 
	\caption[Illustration of the probability based cluster mass prediction method.]{Illustration of the probability based cluster mass prediction method. \emph{Left}: The two dimensional posterior probability distribution of LOSVD and redshift used to determine the correct cluster mass. The pink and green rectangles show the locations of two example galaxy clusters used to create the conditional probability distribution of the mass, $P(\log\ M_{200c}|z, \mathrm{Log}\, \sigma)$. \emph{Right}: The conditional probability distribution of the cluster mass for the two example galaxies. See text for a complete description.}
	\label{fig: probability corner} 
\end{figure*}

In this method, the cluster masses are predicted using the method illustrated in Figure~\ref{fig: probability corner}. The left panel shows the two dimensional (joint probability) projections of the posterior probability distributions of the feature training data. The conditional probability of the cluster mass $P(\log\ M_{200c}|\vec{x}= \{ x_1,x_2,...\})$ (shown in the right panel) is determined by selecting a region in the joint probability distribution. For example, using the LOSVD and redshift features we create $P(\log\ M_{200c}|\vec{x})$ for two test galaxies, shown by the green and pink boxes in the left panel of Figure~\ref{fig: probability corner}. These example galaxies have features $\vec{x} = \{\sigma=200\, \kms,z=0.16\}$ and $\vec{x} = \{\sigma=400\, \kms,z=0.36\}$. We select all galaxies in our training sample with similar features and create the conditional probability distributions shown in the right panel.

For the clusters making up the \emph{test} sample the mass is unknown (it is what we are trying to predict) but the other features are known. To determine the mass probability distribution of a test cluster, $P(\log\ M_{200c})$, we combine the conditional probability distribution, $P(\log\ M_{200c}|\vec{x})$, created previously with the probability distribution of $\sigma$, the LOSVD, through Equation~\ref{eq: Pm}.
\begin{equation}\label{eq: Pm}
	P(\log\ M_{200c}) = \int P(\log\ M_{200c}|\vec{x})\ P(\log\ \sigma)\ d\log\ \sigma\ P(z)\ dz
\end{equation}
The expected mass is determined by calculating the first moment of the probability density. This becomes our ``predicted'' cluster mass, $M_{pred}$.
\begin{equation}\label{eq: expected mass}
	\log\ M_{pred}= \int \log\ M_{200c}\ P(\log\ M_{200c})\ d\log\ M_{200c}
\end{equation}
The confidence interval associated with this prediction can be estimated two ways. First, by calculating the second moment of the probability density through
\begin{equation}\label{eq: variance}
	\mathrm{Var} = \int (\log\ M_{200c} - \log\ M_{pred})^2\ P(\log\ M_{200c})\ d\log\ M_{200c}
\end{equation}
or by drawing many samples from $P(\log\ M_{200c})$ and calculating the values at the 16th and 84th percentile. In practice we find that both methods produce similar results for a large number of trials. Therefore, we quote predicted masses as the most probable mass given by Equation~\ref{eq: expected mass} and associated 68\% error estimated through the square root of Equation~\ref{eq: variance}.

\subsubsection{Machine Learning Based}\label{sec:machine learning method}
The cluster mass estimation in this section relies on a ML technique known as an ensemble method, where many estimators are created by a single learning method with the goal of improved generalization and robustness compared to a single estimation. Ensemble methods \hbox{\citeeg{Caruana2006}} come in two general flavors. Averaging methods average (hence the name) the estimators to produce a single prediction. Boosting estimators build estimates sequentially by attempting to address poor performing estimators in each previous step, hence ``boosting'' the predictive power.

@@

As discussed in Section \ref{sec: observations}, the observational constraints place limits on the total number of clusters member galaxies expected to be recovered. Knowing these limits will provide important information for potential future follow up or Targeted observations. We recover 14,189 clusters with Targeted observations and 1,760 clusters with Survey observations, where we require a detection of $N_{obs} >=5$ galaxies for a cluster to be detected. It is worth noting that we are using galaxies which are a priori known to belong to each cluster. This is not possible to do when analyzing real data, but this will be investigated in a follow up paper, Boada \etal\ (in Preparation).

Figure \ref{fig: recovery} shows the recovery fraction of member galaxies, the number of observed galaxies divided by the number of actual galaxies ($N_{obs}/N_{True}$), as a function of both redshift and cluster mass. As expected, the Targeted observing strategy where the individual clusters are targeted through several dithers to ensure near complete coverage, performs significantly better than the Survey observing strategy across all redshifts and cluster masses. By definition, with Perfect knowledge, recovery fraction would be unity across all redshifts and cluster masses where clusters exist.

@@

\subsection{Mass estimates}
\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/Probcomparison.pdf} 
	\caption[Probability based cluster mass predictions.]{Mass predictions for the power law scaling relation (Equation~\ref{eq:power law}) and the probability based technique with different input features as a function of true cluster mass. The bottom row of panels shows the fractional error (Equation~\ref{eq: fractional error}) also as a function of true cluster mass. The solid black line shows the 1:1 relation. The orange, dash-dotted line is the median predicted mass for perfect observations. The solid, purple line is the median predicted mass for the Targeted observing, and the green, dashed line is the median recovered mass for the HETDEX-like observations. The shaded regions represent the 68\% scatter around the median values.}
	\label{fig:Probability comparison} 
\end{figure*}

@@

\subsection{Impact of Training Sample Cosmology}
All simulations (including that used for Buzzard) use specific values for cosmological parameters. When using simulation data to train ML methods, we incorporate all of those assumptions into the learned feature associations. One could imagine that the specific values of the cosmological parameters in the training sample could bias the ML results when applied to data (real or simulated) created from an unknown true set of cosmological parameters. In addition, each simulation realizes galaxy populations in slightly different ways. The details of the simulation affect which galaxies are observed or not, and thus influences our ability to recover the desired cluster properties.
